---
title: "Initial Korean covid 19"
author: "Cienna Baker-Salgado"
date: "Fall 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

#Methods
Sex is categorical, gender is categorical, the dying result is categorical. CAT~CAT helper file is used since our groups are classified as categorical.

$H_0$ Dying is not independent of sex.
$H_A$ Dying is dependent of sex.




## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The display on the bar chart shows that men have died at higher rates compared to women. 


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Among the people who died, 64% of the population was men and the other 36% were women.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value we found during the chi-squared test was 0.001 so we reject the null hypothesis. We can conclude that dying is dependent of sex. This leaves men 2.3 times as likely to die of COVID-19 compared to women in Korea.

## Conclusions

Seeing out data, we can conclude that deaths based on COVID from Korea are dependent on sex. The Fisher's Exact Test, we find out that Korean men are 2.3 times as likely to die from COVID than women. Our p-value found from the chi-squared test is 0.001, which is less than 0.05. This leads to the rejection of the null hypothesis and the assumption of dying being dependent on sex.